Practical Optional Types for Clojure

### 2015-02-28 related-work.tex

\paragraph{Multimethods} 
\citet{MS02} describe Dubious, a simple statically typed core language including multimethods that
dispatch on the type of its arguments. 
In Dubious methods declare the types of arguments that they expect which corresponds to
exclusively using \clj{class} as the dispatch function in Typed Clojure.

@@

\paragraph{Record Types} 
OCaml-style extensible record types (eg. \citet{Wan89}, \citet{CM91}, \citet{HP91})
are defined polymorphically in terms of row variables.
HMaps in Typed Clojure are instead designed to be used with subtyping.

@@

>>>>

% Cite a few of the early papers here.
%http://www.cs.washington.edu/research/projects/cecil/www/pubs/
\paragraph{Multimethods} 
\citeauthor{MS02} and collaborators present a sequence of
systems~\cite{MS02, cecil, more} with statically-typed multimethods
and modular type checking.  In contrast to Typed Clojure, in these
system methods declare the types of arguments that they expect which
corresponds to exclusively using \clj{class} as the dispatch function
in Typed Clojure. However, Typed Clojure does not attempt to rule out
failed dispatches at runtime.

@@

\paragraph{Record Types} Row polymorphism~\cite{Wan89,CM91,HP91}, used
in systems such as the OCaml object system, provides many of the
features of HMap types, but defined using universally-quantified row
variables. HMaps in Typed Clojure are instead designed to be used with
subtyping, but nonetheless provide similar expressiveness, including
the ability to require presence and absence of certain keys.

@@

The integration of completeness information, crucial for many examples
in Typed Clojure, is not provided by any of these systems.

\paragraph{Java Interoperability in Statically Typed Languages}
Scala~\cite{OCD+} has nullable references for compatibility with Java.
Programmers must manually check for
\java{null} as in Java to avoid null-pointer exceptions.
